\mytitle

### 2021-12-12 paper.tex
rename symbols to avoid confusion
reduce number of coefficients

We define the real dilogarithm $\Li(x)$ for all $x\in\mathbb{R}$ as
the real part of its complex continuation.  The real dilogarithm can
be mapped onto the interval $[0,1/2]$ by using the following
identities:
%
\begin{widetext}%
\begin{align}%
  \Li(x) &= \Li\left(\frac{1}{1-x}\right) - \frac{\pi^2}{6} + \ln(1-x)\left[\frac{1}{2}\ln(1-x) - \ln(-x)\right], && (x<-1), \\
  \Li(x) &= -\Li\left(\frac{x}{x-1}\right) - \frac{1}{2}\ln^2(1-x), && (-1<x<0), \\
  \Li(x) &= -\Li(1-x) + \frac{\pi^2}{6} - \ln(x) \ln(1-x), && (\frac{1}{2}\leq x<1), \\
  \Li(x) &= \Li\left(1-\frac{1}{x}\right) + \frac{\pi^2}{6} - \ln(x)\left[\ln\left(1-\frac{1}{x}\right) + \frac{1}{2}\ln(x)\right], && (1<x<2), \\
  \Li(x) &= -\Li\left(\frac{1}{x}\right) + \frac{\pi^2}{3} - \frac{1}{2} \ln^2(x), && (x\geq 2),%
\end{align}%
\end{widetext}%
%
as well as the special values $\Li(0)=0$, $\Li(1)=\pi^2/6$ and
$\Li(-1)=-\pi^2/12$.  Within the interval $[0,1/2]$ the real
dilogarithm has the series expansion \eqref{eq:series}, which can be
written as
%
\begin{align}
  \Li(x) = x \sum_{k=0} \frac{x^{k}}{(k+1)^2}.
  \label{eq:series2}
\end{align}
%
We approximate the sum on the r.h.s.\ of Eq.~\eqref{eq:series2} using
an economized Padé approximant on the interval $[0,1/2]$ using the
\texttt{EconomizedRationalApproximation} from Wolfram/Mathematica
\cite{mathematica},
%
\begin{align}
  \sum_{k=0} \frac{x^{k}}{(k+1)^2} =
  \frac{\sum_{i=0}^7 a_iy^i}{\sum_{j=0}^7 b_jy^j},
\end{align}
%
where $y=1/4$ is the midpoint of the interval $[0,1/2]$.  The
coefficients $a_i$ and $b_j$ ($i,j=0,\ldots,7$) are listed in
\tabref{tab:coeffs}.

In \tabref{tab:runtime} the runtime of different C implementations is
compared on two different CPU architectures.  On each architecture the
implentations are compiled with \texttt{g++} 10.2.1 with optimization
\texttt{-O2}.  The shown runtime denotes the total wall clock time for
the calculation of $\Li(x)$ for $10^6$ uniformly distributed random
values $x\in[-5,5]$.
%
\begin{table}[tb]
  \centering
  \caption{Runtime in seconds for the calculation of $\Li(x)$ for
    different implementations on different CPU architectures.}
  \begin{tabular}{lcc}
    \toprule
    Implementation & i5-8265U & i7-5600U \\
    \midrule
    this paper & $0.022$ & $0.027$ \\
    Morris \cite{morris} & $0.022$ & $0.027$ \\
    Cephes \cite{cephes} & $0.028$ & $0.034$ \\
    Algorithm 327 \cite{koelbigDilog} & $0.031$ & $0.035$ \\
    Chebyshev \cite{luke} & $0.058$ & $0.060$ \\
    Algorithm 490 \cite{ginsberg} & $0.059$ & $0.071$ \\
    GSL \cite{gsl} & & $0.155$ \\
    \bottomrule
  \end{tabular}
  \label{tab:runtime}
\end{table}
%
We find that the implementation presented in this paper is fastest
among the ones shown in the table.  The next-to fastest implementation
is provided by Cephes, which is only $\sim 20\%$ slower.  The short
runtime of Cephes stems from the fact that Cephes also uses a rational
function approximation, where the numerator and the denominator can be
evaluated in parallel.  However, Cephes evaluates the polynomials in
the numerator and denominator using Horner's scheme, which is purely
sequential.  Furthermore, for $x>1$ Cephes calculates $\ln(x)$ twice,
which leads to a further slowdown.

>>>>

We define the real dilogarithm $\Li(x)$ for all $x\in\mathbb{R}$ as
the real part of its complex continuation.  The real dilogarithm can
be mapped onto the interval $[0,1/2]$ by using the following
identities:
%
\begin{widetext}%
\begin{align}%
  \Li(x) &= \Li\left(\frac{1}{1-x}\right) - \frac{\pi^2}{6} + \ln(1-x)\left[\frac{1}{2}\ln(1-x) - \ln(-x)\right], && (x<-1), \\
  \Li(x) &= -\Li\left(\frac{x}{x-1}\right) - \frac{1}{2}\ln^2(1-x), && (-1<x<0), \\
  \Li(x) &= -\Li(1-x) + \frac{\pi^2}{6} - \ln(x) \ln(1-x), && (\frac{1}{2}\leq x<1), \\
  \Li(x) &= \Li\left(1-\frac{1}{x}\right) + \frac{\pi^2}{6} - \ln(x)\left[\ln\left(1-\frac{1}{x}\right) + \frac{1}{2}\ln(x)\right], && (1<x<2), \\
  \Li(x) &= -\Li\left(\frac{1}{x}\right) + \frac{\pi^2}{3} - \frac{1}{2} \ln^2(x), && (x\geq 2),%
\end{align}%
\end{widetext}%
%
as well as the special values $\Li(0)=0$, $\Li(1)=\pi^2/6$ and
$\Li(-1)=-\pi^2/12$.  Within the interval $[0,1/2]$ the real
dilogarithm has the series expansion \eqref{eq:series}, which can be
written as
%
\begin{align}
  \Li(x) = x \sum_{k=0} \frac{x^{k}}{(k+1)^2}.
  \label{eq:series2}
\end{align}
%
We approximate the sum on the r.h.s.\ of Eq.~\eqref{eq:series2} using
a rational minimax approximant on the interval $[0,1/2]$ using the
\texttt{MiniMaxApproximation} from Wolfram/Mathematica
\cite{mathematica},
%
\begin{align}
  \sum_{k=0} \frac{x^{k}}{(k+1)^2} =
  \frac{\sum_{k=0}^5 p_kx^k}{\sum_{k=0}^6 q_kx^k}.
  \label{eq:minimax}
\end{align}
%
The coefficients $p_k$ and $q_k$ are listed in \tabref{tab:coeffs}.
%
\begin{table}[t]
  \centering
  \caption{Coefficients of the numerator and denominator polynomials
    for the minimax approximant \eqref{eq:minimax}.}
  \begin{tabular}{lr}
    \toprule
    $p_0$ & $ 0.9999999999999999502\cdot 10^{+0}$ \\
    $p_1$ & $-2.6883926818565423430\cdot 10^{+0}$ \\
    $p_2$ & $ 2.6477222699473109692\cdot 10^{+0}$ \\
    $p_3$ & $-1.1538559607887416355\cdot 10^{+0}$ \\
    $p_4$ & $ 2.0886077795020607837\cdot 10^{-1}$ \\
    $p_5$ & $-1.0859777134152463084\cdot 10^{-2}$ \\
    $q_0$ & $ 1.0000000000000000000\cdot 10^{+0}$ \\
    $q_1$ & $-2.9383926818565635485\cdot 10^{+0}$ \\
    $q_2$ & $ 3.2712093293018635389\cdot 10^{+0}$ \\
    $q_3$ & $-1.7076702173954289421\cdot 10^{+0}$ \\
    $q_4$ & $ 4.1596017228400603836\cdot 10^{-1}$ \\
    $q_5$ & $-3.9801343754084482956\cdot 10^{-2}$ \\
    $q_6$ & $ 8.2743668974466659035\cdot 10^{-4}$ \\
    \bottomrule
  \end{tabular}
  \label{tab:coeffs}
\end{table}

In \tabref{tab:runtime} the runtime of different C implementations is
compared on two different CPU architectures.  On each architecture the
implentations are compiled with \texttt{g++} 10.2.1 with optimization
\texttt{-O2}.  The shown runtime denotes the total wall clock time for
the calculation of $\Li(x)$ for $10^6$ uniformly distributed random
values $x\in[-5,5]$.
%
\begin{table}[tb]
  \centering
  \caption{Runtime in seconds for the calculation of $\Li(x)$ for
    different implementations on different CPU architectures.}
  \begin{tabular}{lcc}
    \toprule
    Implementation & i5-8265U & i7-5600U \\
    \midrule
    this paper & $0.022$ & $0.026$ \\
    Morris (DILOG 0011) \cite{morris} & $0.022$ & $0.026$ \\
    Cephes \cite{cephes} & $0.028$ & $0.034$ \\
    Algorithm 327 \cite{koelbigDilog} & $0.031$ & $0.035$ \\
    Chebyshev \cite{luke} & $0.058$ & $0.060$ \\
    Algorithm 490 \cite{ginsberg} & $0.059$ & $0.071$ \\
    GSL \cite{gsl} & & $0.155$ \\
    \bottomrule
  \end{tabular}
  \label{tab:runtime}
\end{table}
%
We find that the implementation presented in this paper is fastest
among the ones shown in the table.  The next-to fastest implementation
is provided by Cephes, which is only $\sim 20\%$ slower.  The short
runtime of Cephes stems from the fact that Cephes also uses a rational
function approximation, where the numerator and the denominator can be
evaluated in parallel.  However, Cephes evaluates the polynomials in
the numerator and denominator using Horner's scheme, which is purely
sequential.  Furthermore, for $x>1$ Cephes calculates $\ln(x)$ twice,
which leads to a further slowdown.
